
# 2- DATA

```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts/setup.R"))
```

## A- Sources

We managed to gather raw data from 4 different sources:

1. **WorldBank** statistics and datasets
2. **Heritage.org** - World Economic Forum statistics
3. **United Nation Development Program** - Human Development Reports
4. **Socio-Economic Data and Application Center** - NASA and Columbia.edu research


## B- Description

Our gathered data consisted of the following:

### Social, Envrionmental and Governmental Indicators

- Traditional Social Indicators:

1. Urbanization **"Urban"** - WB
*(% of urban landscape in a country Vs Total Land)* 

2. Female Labour Force Participation **"FemLab"** - WB
*(% of Labour Force)* 

3. Aging Population **"AgingPop"** - WB
*(% of pop above 65 years old) "* 

- ESG Scoring Indexes:

4. Education Index **"EduI"** - UNDP
*(model showing score of Edu quality)* 

5. Life Expectancy Index **"LEI"** - UNDP
*(model showcasing health)* 

6. Income Index **"IncI"** - UNDP
*(model showing income level/quality)* 

7. Human Development Index **"HDI"** - UNDP
*(model that includes life expectancy, education and knowledge as well as GNI factors)* 

8. Economic Freedom Index **"EFI"** - Heritage / WEF
*(model that factors in Judicial aspects, business, investment, trade, labour and monetary freedom scores; fiscality and government spending and integrity)* 

9. Environmental Performance Index **"EPI"**- SEDAC
*(model that factors in Environmental health: Waste Mngt, Water poll, air quality, heavy metals; also Ecosystem Vitality: Climate change, biodiversity, services, fisheries, pollution and agriculture)* 

**EPI Modeling**
NOTE!: We recreated the Environmental Performance index model from scratch by using all the components' raw data and taking into consideration exact weighting schemes and allocations.  We however modified a minor detail; by removing the weight from Fisheries and reallocating it to other aspects , especially Climate Change, we managed to include many more countries with viable data in earlier years (approx twice more countries included from the re-weighting adjustment we did). 
Climate is bigger issue and a more serious topic compared to Fisheries all over the world and we believed that this re-weighting adjustment focuses better on Environmental Performance in our context compared to the traditional model.
It took a lot of work but we made sure the indicator and its adjustments better suited our research needs. 
*The model has been adjusted and calibrated on Excel as linked sheets and forumlas are way easier and more practical; doing it on R would be too time consuming and would not provide any added value for our project since it is outside the scope of our course.The excel file is in the folder shared*


### Economic Indicators

10. Unemployment Rate *in % Yearly* **"Unemploy"** - WB

11. GDP per Capita *in current USD* **"GDPpcap"** - WB

12. GDP growth *% annual* **"GDPg"**- WB

13. Foreign Direct Investment *in current USD in millions* **"FDIusd"** - WB

14. Foreign Direct Investment *as % of GDP* **"FDIGDP"** - WB

### Data, Periodicity and Frequency

Data was gathered on annual basis given the lack of shorter frequency for all scoring indexes (EPI, EFI .....). 
Time period was starting form 1990 until 2019 (Also given the data limitations of most indicators)
For the EFI and EPI data before 1995 was non existant. Data for LEI, IncI, EduI and HDI was missing for 2019.
Also we had many missing NA values in some of our data, we had some very minor linear interpolation done on Excel for big countries but we kept the NA values as is in order not to manipulate any results.


## C- Data Wrangling

These are the following steps used to import and wrangle our data:

1- Using the excel "Dataset.raw" in our data files containing different sheets of different indicators for the countries over the years we used a function after reading the data to import every sheet by itself and upload them as separate tibbles in our Rstudio environment; we also made sure to keep the sheet names in the excel are the same as the related tibble names.


```{r fig.show = "hide", warning=FALSE, results='hide'}
sheets <- readxl::excel_sheets(here::here("data/Dataset.raw.xlsx"))
lst <- lapply(sheets, function(sheet) 
  readxl::read_excel(here::here("data/Dataset.raw.xlsx"), sheet = sheet))

names(lst) <- sheets

list2env(lst, envir = .GlobalEnv)

rm(`Country Table`)
```

Most of our data sets of Indicators contain between 120 and 200 countries , 25 to 30 years of data.

### Sample Indicator Dataset: HDI

A small sample of our Dataset related to one of our Indicators "Human Development Index  - HDI" is shown below

```{r}

kbl(HDI[1:15, 1:6], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) 

```


2- After noticing that some values were not imported as numerics we had to convert all columns with non-characted values into numerics in order to conduct analysis on the data later

```{r warning = FALSE}

cols <- c("1990", "1991","1992","1993","1994","1995","1996","1997","1998","1999", "2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010",
          "2011","2012","2013","2014","2015","2016","2017","2018","2019")

FemLab[,cols] <- sapply(FemLab[,cols],as.numeric)
AgingPop[,cols] <- sapply(AgingPop[,cols],as.numeric)
Urban[,cols] <- sapply(Urban[,cols],as.numeric)
Unemploy[,cols] <- sapply(Unemploy[,cols],as.numeric)
FDIGDP[,cols] <- sapply(FDIGDP[,cols],as.numeric)
FDIusd[,cols] <- sapply(FDIusd[,cols],as.numeric)
GDPg[,cols] <- sapply(GDPg[,cols],as.numeric)
GDPpcap[,cols] <- sapply(GDPpcap[,cols],as.numeric)
EFI[,cols] <- sapply(EFI[,cols],as.numeric)
LEI[,cols] <- sapply(LEI[,cols],as.numeric)
HDI[,cols] <- sapply(HDI[,cols],as.numeric)
EduI[,cols] <- sapply(EduI[,cols],as.numeric)
IncI[,cols] <- sapply(IncI[,cols],as.numeric)
EPI[,cols] <- sapply(EPI[,cols],as.numeric)

```

3- After considering the best way to combine and plot the data visually or use it for quantitative analysis by groups we decided to create a pivoted version of each data set in order to ultimately join all of them together. We Used 1 column for the Years another for the Indicator Value.

```{r}

AgingPop <- AgingPop %>%  
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "AgingPop")
FemLab <-  FemLab %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "FemLab")
Urban <-Urban %>% 
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "Urban")
Unemploy<-Unemploy %>% 
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "Unemploy")
FDIusd<- FDIusd %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "FDIusd")
FDIGDP <- FDIGDP %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "FDIGDP")
GDPg<- GDPg %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "GDPg")
GDPpcap <- GDPpcap %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "GDPpcap")
EFI <- EFI %>% 
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "EFI")
EPI <- EPI %>% 
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "EPI")
EduI <- EduI %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "EduI")
IncI <- IncI %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "IncI")
LEI <- LEI %>%
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "LEI")
HDI <- HDI %>% 
  pivot_longer(c('1990':'2019'), names_to= "Year", values_to = "HDI")

```

### Sample Pivoted Table: GDP per Capita

A sample of the pivoted table for GPD per Capita "GDPpcap" is shown below

```{r}
kbl(GDPpcap[1:20,], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")
```

4- We used inner_join to combine all the seperate indicator datasets into one Full Table in order to study the data more practically

### The Full Data Table

A summary table including only countries that have data in each and every Indicator

```{r}

Full.Table <- Urban %>% 
  inner_join(AgingPop[,3:5], by = c("Country","Year")) %>%
  inner_join(FemLab[,3:5], by = c("Country","Year")) %>%
  inner_join(Unemploy[,3:5], by = c("Country","Year")) %>%
  inner_join(GDPg[,3:5], by = c("Country","Year")) %>%
  inner_join(GDPpcap[,3:5], by = c("Country","Year")) %>%
  inner_join(FDIGDP[,3:5], by = c("Country","Year")) %>%
  inner_join(FDIusd[,3:5], by = c("Country","Year")) %>%
  inner_join(HDI[,3:5], by = c("Country","Year")) %>%
  inner_join(LEI[,3:5], by = c("Country","Year")) %>%
  inner_join(IncI[,3:5], by = c("Country","Year")) %>%
  inner_join(EduI[,3:5], by = c("Country","Year")) %>%
  inner_join(EPI[,3:5], by = c("Country","Year")) %>%
  inner_join(EFI[,3:5], by = c("Country","Year"))
  
colnames(Full.Table)[1] <- ("Income.Group")

kbl(Full.Table, booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")

```

## D- Data Overview

The below is summary a table to show a small quantitative preview of our data in our indicators.

```{r}
summary(Full.Table)[,5:18]
```
We observe the NA values which were already discussed before importing the data. At a first glance we can also see the huge variance and gaps in our FDI data and our GDP per cap. 



```{r}

write.csv(Full.Table, here::here("data/Full.Table"), row.names = FALSE)

```


