
# 3- EXPLORATORY  DATA  ANALYSIS

```{r, echo = FALSE, message = FALSE, warning=FALSE, cache=TRUE}
source(here::here("scripts/setup.R"))
```

***In this section we will be:***

- Identifying the main indicators for EDA
- Mapping out the basis of our structure and analysis
- Creating Summary Tables to better visualize and plot relationships
- Plotting some univariate and multivariate visuals


```{r}
Full.Table <- read.csv(here::here("data/Full.Table"))
```

## A- The ESG Composite

Creating our own ESG Index with the following weights:

- EPI 30%
- EFI 30%
- LEI IncI EduI  40%/3 each

We also mutate and add this to our existing Full table to have our complete table

```{r}
Complete.Table <- Full.Table %>%
mutate(ESGcomp = 0.3*EPI + 0.3*EFI + 0.4/3*LEI + 0.4/3*IncI + 0.4/3*EduI)

kbl(Complete.Table[140:160,c(1:4,15:19)], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")
```


## B- Scalability and Comparison

A major issue with our Data , especially the values in USD (like GDP/Capita and FDI), is that the values are huge compared to our scoring indexes of scale 100% and our traditional social indicators which represent percentages as well.

**Scaling** the data in a range would misrepresent that variations visually and statistically. 

We decided to import a rebased version of the same Complete.Table dataset with rebasing starting from 100 at year 1990 (taking the first available datapoint and tracing back in case there is an NA value at 1990).The data set is called Complete Scaled and has rebased value done through excel for simplicity as well and so as not to overcrowd our R studio with inefficient and out of scope codes. 

**Rebased Complete Table**
<div style="margin-bottom:100px;">
```{r warning = FALSE}
Rebased.Table <- read_excel(here::here("data/Complete.Scaled.xlsx"))


cols2<- c("Urbanr","AgingPopr","FemLabr","Unemployr","GDPgr","GDPpcapr",
          "FDIGDPr","FDIusdr","HDIr","LEIr","IncIr","EduIr","EPIr","EFIr","ESGcompr")

Rebased.Table[,cols2] <- sapply(Rebased.Table[,cols2],as.numeric)

kbl(Rebased.Table, booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")
```
</div>


We also transpose the Rebased Table by pivotting into countries as columns with Indicator values and Years as rows for different visibility and easier plotting

```{r}

Rebased.Pivot <- Rebased.Table %>%
pivot_longer(c("Urbanr":"ESGcompr"),names_to = "Indicator", values_to= "Value")

kbl((Rebased.Pivot%>%arrange(Country, Year,Indicator))[1:20,], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")
```


## C- Data Exploration

### Exploring the G9

To visually test our Data sets we decided to take a random sample of countries and chose the G9 so we can observe different elements of our data.

**Rebased G9 Table**

```{r}
G9.Table <- Rebased.Pivot %>%
filter(Country == "United States" | Country == "United Kingdom" | Country == "Canada" |
Country == "Russia" | Country == "China" |Country == "Italy" | Country == "Japan" |
Country == "France" | Country == "Germany")

kbl(G9.Table, booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")

```

### Testing the ESG Composite

A plot showing the time series of the ESG composite of the G9 countries

```{r}
ggplotly(ggplot(G9.Table %>%
  filter( Indicator == "ESGcompr"),aes(x=Year, y= Value, group= Country ))+
  geom_line(aes(color= Country))+
  theme_gray(base_size = 16, base_family = "Times")+ 
  ggtitle("ESG Index for G9 countries") +
  ylab("ESG Score - Rebased"), height = 400, width = 8)
```

*Observation: A clear positive trend over the years for most countries*


### Comparing ESG, FDI and GDP

We are attempting to plot the rebased GDP per capita, FDI in usd and ESG score for the G9 countries just to spot the magnitude of the differences of these indicstors and their trends over the past 30 years

```{r ,fig.width= 9}
ggplotly(ggplot(G9.Table %>%
  filter( Indicator == "ESGcompr" | Indicator == "GDPpcapr"| Indicator == "FDIusdr") 
  ,aes(x= Year, y= Value, group = Indicator ))+
  geom_line(aes(color= Indicator))+
  scale_color_manual(values = c("darkred", "steelblue","deeppink","grey45","grey0"))+
  facet_wrap_paginate(
    ~Country,
    nrow = 3,
    ncol = 3,
    scales = "fixed",
    shrink = TRUE,
    labeller = "label_value",
    as.table = TRUE,
    switch = NULL,
    drop = TRUE,
    dir = "h",
    strip.position = "top",
    page = 1
  )+
  theme_gray(base_size = 10, base_family = "Times")+
  ggtitle("Social Indices for G9 countries" ), height = 450
)
```

*Observation: It is clear that the FDI is too volatile and uncorrelated with the data to be further used with our analysis*


### Visualising our ESG Score Indexes

A time series showing the countries' EFI, EPI, IncI, EduI and LEI indexes from 1990 till date.


```{r, fig.width= 9}
ggplotly( ggplot(G9.Table %>%
  filter( Indicator == "IncIr" | Indicator == "EPIr"| Indicator == "EFIr"| Indicator == "EduIr"| Indicator == "LEIr")  ,aes(x= Year, y= Value, group = Indicator ))+
  geom_line(aes(color= Indicator))+
  scale_color_manual(values = c("darkred", "steelblue","deeppink","grey45","grey0"))+
  facet_wrap_paginate(
    ~Country,
    nrow = 3,
    ncol = 3,
    scales = "fixed",
    shrink = TRUE,
    labeller = "label_value",
    as.table = TRUE,
    switch = NULL,
    drop = TRUE,
    dir = "h",
    strip.position = "top",
    page = 1
  )+
  theme_gray(base_size = 10, base_family = "Times")+
  ggtitle("Social Indices for G9 countries" ) , height= 450
)
```
*Observation: EFI and EduI don't seem as correlated as other indicators*

### The G9 Correlation Matrices

A more quantitative visual assessment showing a correlation matrices of the G9 countries between all indicators across time


```{r, warning=FALSE, out.width="100%"}
CorA <- split(( G9.Table%>%
  pivot_wider(names_from = "Indicator", values_from = "Value" ) %>%
  droplevels())[,5:19], ( G9.Table%>%
  pivot_wider(names_from = "Indicator", values_from = "Value" ) %>%
  droplevels())$Country)

# The below function should be selected all together and loaded for optimal results
par(mfrow=c(3,3)) 
for(i in seq_along(CorA)){
  corrplot((rcorr(as.matrix(CorA[[i]])))$r, type="upper",tl.col="black", tl.cex=0.7,tl.srt=70,   col=colorRampPalette(c("red","white","blue"))(40),
           p.mat = (rcorr(as.matrix(CorA[[i]])))$P, insig = "blank",sig.level = 0.01)
  mtext(paste(names(CorA)[i]),line=2,side=3)}
```        

*Observation: Concentration of correlation around the Score Indexes and GDP per Capita compared to a lower correlations with traditional indicators, although Urbanisation and to some extenet Female Labour seem to correlate with Indexes*


### Traditional Social Indicators

Assessing the relationship between Female Labour, Urbanisation, Aging Population Vs the GDP per Capita for the G9 countries in time series

```{r}
ggplotly( ggplot(G9.Table %>%
  filter( Indicator == "FemLabr" | Indicator == "Urbanr"| Indicator == "AgingPopr"| Indicator == "Unemployr"| Indicator == "GDPpcapr")  ,aes(x= Year, y= Value, group = Indicator ))+
  geom_line(aes(color= Indicator))+
  scale_color_manual(values = c("darkred", "steelblue","deeppink","grey45","grey0"))+
  facet_wrap_paginate(
    ~Country,
    nrow = 3,
    ncol = 3,
    scales = "fixed",
    shrink = TRUE,
    labeller = "label_value",
    as.table = TRUE,
    switch = NULL,
    drop = TRUE,
    dir = "h",
    strip.position = "top",
    page = 1
  )+
  theme_gray(base_size = 10, base_family = "Times")+
  ggtitle("Social Indicators for G9 countries" ) ,height = 400
)

```
*Observation: No obvious distinct relationship observed for these countries*

### ESG Indexes Vs Social Indicators

Assessing the relationship between traditional social and other indicators Vs the Scoring ESG indexes for the G9 to try and spot significant relationships

```{r}
ggplotly( ggplot(G9.Table %>%
  filter( Indicator == "FemLabr" | Indicator == "Urbanr"| Indicator == "AgingPopr"| Indicator == "EPIr"| Indicator == "EFIr"| Indicator == "HDIr")  ,aes(x= Year, y= Value, group = Indicator ))+
  geom_line(aes(color= Indicator))+
  scale_color_manual(values = c("darkred", "steelblue","deeppink","grey45","grey0","green", "yellow"))+
  facet_wrap_paginate(
    ~Country,
    nrow = 3,
    ncol = 3,
    scales = "fixed",
    shrink = TRUE,
    labeller = "label_value",
    as.table = TRUE,
    switch = NULL,
    drop = TRUE,
    dir = "h",
    strip.position = "top",
    page = 1
  )+
  theme_gray(base_size = 10, base_family = "Times")+
  ggtitle("ESG Vs Traditional Social Indicators" ) ,height = 400
)
```
*Observation: No obvious relationship can be concluded but we cannot rule out that these indicators might have a sort of correlation especialy for the developed Economies*


### The Scatter Plot: 20 Years

Create a summary table for all indicators (non-rebased) for the year 2018, 2008 and 1998 and dropping all countries with any NA value.
The purpose is to showcase how in time period of 10 years Scoring Indexes for different countries, regions and Income Groups are changing.

**Data Table for: 1998-2008-2018**

```{r}

Data.98.08.18 <- Complete.Table %>%
filter(Year == 2018 | Year == 2008 |Year == 1998) %>%
na.omit()


kbl(Data.98.08.18[20:50,], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")
```



We will now plot a Y/X scaterplot showing EPI vs EFI for latest of the 3 above years for all relevant countries. The scaterplot shows each point sized by GDP per capita (non-rebased) with colors differing based on region.

**EPI Vs EFI - 1998 Vs 2008 Vs 2018**

```{r dpi= 300, fig.width= 8}


ggplotly(ggplot(Data.98.08.18, aes(x= EPI, y=EFI,group=Country))+
  geom_point(aes(color=Region,size=GDPpcap, shape = Income.Group))+
  facet_wrap( ~Year,  nrow = 1, ncol = 3)+
  ggtitle("EPI Vs EFI Over 20 years period" ) +
  xlab("EPI")+
  ylab("EFI")+
  theme(legend.text = element_text(color = "grey", size = 5), legend.position = "left"), legend.position = "left", height= 400)
   
```

*Observation: Clearly seen that Europe & Central Asia Have the highest consistent EFI and EPI scores. North America shows a rise in score over time and East Asia a more significant rise. Also observed is the increase of GDP per capita for these regions over time, an association of GDP and these indicators might proven in our later analysis*


###  Exploring Country Categories

We created Summary Tables Taking the Average of all countries per year for each indicator based on grouping of both Region and Income Group. All values used are rebased for better visualization.

The ideas is to see whether the simple arithmetic average score or value of our indicators for each income group and each region show a relationship over time.


**Average Indicator Value for Income Groups**

We created a table grouped by income groups that takes the average, for each indicator (per year), of all countries in that group.

```{r warning = FALSE}

IncGrp.Avg <- Rebased.Table %>%
group_by(Income.Group, Year) %>%
summarize_all(funs(mean(.,na.rm =TRUE))) %>%
select(-c(Region,Country)) %>%
  pivot_longer(c('Urbanr':'ESGcompr'), names_to= "Indicator", values_to = "Value")

kbl(IncGrp.Avg[1:20,], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "450px")
```

**Plot of EFI, EPI, HDI and GDP For All Income Groups**

```{r}
ggplotly( ggplot(IncGrp.Avg %>%
  filter( Indicator == "EPIr"| Indicator == "EFIr"| Indicator == "HDIr")  ,aes(x= Year, y= Value))+
  geom_line(aes(color= Indicator))+
  scale_color_manual(values = c("darkred", "steelblue","deeppink","grey45","grey0","green", "yellow"))+
  facet_wrap_paginate(
    ~Income.Group,
    nrow = 2,
    ncol = 2,
    scales = "fixed",
    shrink = TRUE,
    labeller = "label_value",
    as.table = TRUE,
    switch = NULL,
    drop = TRUE,
    dir = "h",
    strip.position = "top",
    page = 1
  )+
  theme_gray(base_size = 10, base_family = "Times")+
  ggtitle("Income Group - Mean Relationships" ), height= 400 
)
```

*Observation: It seems the higher the income group level the stronger the correlation among these indicators*


**Average Indicator Value for Regions**

We did the same average study but now we grouped by Region...

```{r warning = FALSE}
Region.Avg <- Rebased.Table %>%
group_by(Region, Year) %>%
summarize_all(funs(mean(.,na.rm =TRUE))) %>%
select(-c(Country,Income.Group))%>%
  pivot_longer(c('Urbanr':'ESGcompr'), names_to= "Indicator", values_to = "Value")

kbl(Region.Avg[1:20,], booktabs = TRUE) %>%
  kable_classic(full_width = T, html_font = "Cambria", "striped")%>%
  kable_styling(bootstrap_options = c("hover", "condensed")) %>% 
  scroll_box(width = "100%", height = "300px")
```

**Plot of EFI, EPI, HDI and GDP For All Regions**

```{r}
ggplotly( ggplot(Region.Avg %>%
  filter( Indicator == "EPIr"| Indicator == "EFIr"| Indicator == "HDIr")  ,aes(x= Year, y= Value ))+
  geom_line(aes(color= Indicator))+
  scale_color_manual(values = c("darkred", "steelblue","deeppink"))+
  facet_wrap_paginate(
    ~Region,
    nrow = 3,
    ncol = 3,
    scales = "fixed",
    shrink = TRUE,
    labeller = "label_value",
    as.table = TRUE,
    switch = NULL,
    drop = TRUE,
    dir = "h",
    strip.position = "top",
    page = 1
  )+
  theme_gray(base_size = 10, base_family = "Times")+
  ggtitle("Region-Mean Relationships" ), height = 400 
)
```

*Observation: It also seems that Regions where Countries have higher GDP per capita and higher income levels, the relationship among Scoring Indexes is positive and stronger compared to smaller economies' regions*


## D- Preliminary Observations

We can see from the primary quantitative and visual analysis that time itself leads to growth of ESG indices and other non-economic indicators as well as GDP. There exists some visible correlation and positive relationship between ESG and economic indicators over time; but the extent and significance of that correlation is yet to be quantified. 
Some positive correlation between the major G9 countries is observed and we can definitely see the impact of GDP on the score of the ESG factors when observed per year. 
FDI on the other hand, specifically the Net inflows data we derived, does not seem to correlate as clearly as GDP. 
Further in depth analysis will be conducted in the next section that filters for specific countries and indicators to test the relationship of our indicators over time and check whether the relationship is significant.


Regarding our Average Score plots for Income Groups and regions:

- On A Regional Level:

ESG indicators do not seem to correlate , but Developed areas such as Europe and North America show more conformity in scores and trends

- On an Income Group Level:

The Indexes show a stronger relationship between income groups

*Note: We realize that a simple arithmetic average of the Rebased scores is not statistically sound and accurate, hence why we will be focusing our analysis on Representative countries and basing our filter on the biggest economies per each category*




```{r}

write.csv(Complete.Table, here::here("data/Complete.Table"), row.names = FALSE)
write.csv(Rebased.Table, here::here("data/Rebased.Table"), row.names = FALSE)
write.csv(G9.Table, here::here("data/G9.Table"), row.names = FALSE)
write.csv(Data.98.08.18, here::here("data/Data.98.08.18"), row.names = FALSE)
write.csv(Rebased.Pivot, here::here("data/Rebased.Pivot"), row.names = FALSE)
```
